---
title: "Bio4158 - Devoir 3"
author: "Groupe XX - Nom1, Nom2, Nom3"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    number_sections: no
---

Here are the needed packages

```{r}
library(readr)
library(tidyverse)
library(ggplot2)
library(car)
library(performance)
library(see)
library(patchwork)
library(lmtest)
library(pwr)
library(broom)
library(plyr)
```

Loading data into R 

```{r}
setwd("/Users/apple/Desktop/BIO_4158 /BIO 4158 lab/Data sets BIO 4158")
getwd()

climate <- read_csv("climate.csv")

head(climate)
```




Answer the following questions. Total marks = 30.

1. What is the ‘biological’ (scientific) hypothesis? (1 point)

** Climate change is observed to be warming.
There is a variation in changes in temperature across different geographic location, and this pheonomena is causes by the variation in human activities across region **


2. What does this hypothesis predict in the context of her study? (2 points)

** This hypothesis predicts that warming is more prevalent in the Northern area compared to the Southern Area **


3. What is the associated null hypothesis? (2 points)

** H_o: North Area and South Are have the same increase in temparature **


4. To test this hypothesis, the student will need to quantify the extent of any climate change separately for each location. She’s interested not only in the point estimate of the rate of climate change for each location (which she needs for testing her biological hypothesis above), but also in its significance in each location (i.e. she wants to separately test whether there is evidence that climate has changed in any way at each site). Let’s start by conducting a regression analysis to test whether there is evidence of climate change in **Yarmouth**  
    a. What is the statistical null hypothesis relevant to the analysis ofthe Yarmouth data? (2 points)
    
    
   **There is no linear relationship between year and average tempterature in Yarmouth**
   
   b) What is the statistical model in the verbal form ? 
  
  Annual_Average_temperature = Year+ Error
    
    
  c) Provide a relevant plot of the Yarthmouth data
  
  **In this case, we would use the scatter plot**
  
```{r}
    Yarmouth <- climate %>% filter(loc=="Y") %>% arrange(year)

head(Yarmouth)

mygraph <- ggplot(
  data = Yarmouth[!is.na(Yarmouth$ave), ], # source of data
  aes(x = year, y = ave)
)
# plot data points, regression, loess trace
mygraph <- mygraph +
  stat_smooth(method = lm, se = FALSE, color = "green") + # add linear regression, but no SE shading
  stat_smooth(color = "red", se = FALSE) + # add loess
  geom_point() # add data points

mygraph
```
    
    
  d) Fit the model with and shows the R output 
    
```{r}
 RegModel.1 <- lm(ave ~ year, data = Yarmouth)
summary(RegModel.1)
```
  e) State the statistical assumptions for the analysis of Yarmouth data and provide some evidence that you examined them
  
  There are several assumptions that we have to make sure:


- No errors in X measurements
- Relationship between X and Y is linear
- Independence of residuals (no serial aurocorrelation)
- Residuals are normally distributed
- Homoscedasticity of residuals (even spread of residuals on X-axis) 


We will provide the graphs for visualization purpose only. We would not use them to intepret the valdity of our assumption




  **residuals plot**
  
```{r}
par(mfrow = c(2, 2), las = 1)
plot(RegModel.1)
```
  **Linearity plots**
```{r}
check_model(RegModel.1)

```

From here, we will start checking the validity of the assumptions using different types of test 

We use the Breusch-Pagan test examines whether the variability of the residuals is constant with respect to increasing fitted values

```{r}
bptest(RegModel.1)

```
From the output, we can see that the **p-value is 0.625 >0.05**, thus we fail to reject the null hypothesis. **Meaning, the variance is constant**


Next, we will use the the Durbin-Watson test to detect serial autocorrelation in the residuals. Under the assumption of no autocorrelation, the D statistic is 2.

```{r}
dwtest(RegModel.1)
```
From the output, we the **p-value is 0.01735 <0.05**, so we reject the null hypothesis. This means that **there is a autocorrelation** in the residuals 


Thirdly, we will use the the RESET test is a test of the assumption of linearity. If the linearity assumption is met, the RESET statistic will be close to 1.

```{r}
resettest(RegModel.1)

```
From the output, we can see that the RESET point = 7.689, which is way bigger than 1 ,and the p-value is 0.001153 <0.05. Thus, we can reject the Null Hypothesis and conclude that there is no linear relationship. 


Finally, we will use the Shapiro-Wilk normality test on the residual to confirm that the deviation from normality of the residuals is not large


```{r}
shapiro.test(residuals(RegModel.1))

```
  From the output, we can see that the p-value is 0.1807 > 0.05. Thus, the values do not deviate too much from Normal Distribution 


Question 4d 

f) We will show the output of the model 

```{r}
summary(RegModel.1)
```

From the output, we can see that the **p-value is 0.582**. Thus, we fail to reject the null Hypothesis. This means that 
the slope is 0 ($\hat{\beta}$)=0). We can also interpret this as the independent variable **year** have no statistically significant relationship with the dependent variable **average temperature**

4g) Provide a "biological" conclusion concerning climate change in Yarmouth (i.e: qualified statemement about climate change at this site including an estimate of raw effect sie, its precision and whether a trend of this magnitude matters).



Calculation of effective size:

 $$ d = \frac{b}{sb\sqrt{n-k-1}} = \frac{0.002649}{0.004785\sqrt{58-1-1}}=0.00739 $$
 So, **effect size is 0.00739**
 
Precision would be based on the standard error. We can also calculate the 95% confidence interval to show the precision. 
 
 
My biological conclusion is that: The change in temperature in Yarmouth is not a serious issue. 



4h) Provide a statement about your confidence in the statistical conclusion (for example with respect to the violation of certain assumptions, to potential biases created by missing data, to suboptimal experimental design, and/or to power issues). 

The data that we are analyzing comes from the city Yarmouth.

```{r}
head(Yarmouth,6)
```


- First, I want to address whether or not the data the violate assumptions for linear regression model. From several tests being done in part e, we can see that the data violate 2 assumptions (there no linear relationship between independenta variable, and there is an autocorrelation between the residuals). Thus, I am **not confident** that the data would give a good analysis. 

- Secondly, I want to address bias due to missing data (NA values). For this data set Yarmouth, that is not an issue because we do not have NA values in the data set.

- Thirdly, I want to discuss the experimental design of the data. Since the data is collected from the goverment website, there is not enough sufficient information for me to have a conclusion on how the data should be collected. 

- Fourth, I want to calculate the power of the hypotheis test on the estimation of climate change rate in Yarth mouth

We have: $n=58, d= 0.00739, \alpha=0.05$

```{r}
pwr.t.test(n = 58, 
           d = 0.00739, 
           sig.level = 0.05, 
           type = "one.sample")
```

From the output, we can see that the **power= 0.050**, which is very low. From this, we know that there is a high chance we will have type II error in our analysis. 


Overall, From addressing quality of data and power, I conclude that we **should not** make any conclusion from this analysis


  
  
5. Yarmouth is only one of the six locations for which she needs to quantify the extent of any climate change. Plot the temporal trend for temperature separately for each station (2 points), and estimate the rate of change for each station (2 points).


First, we will create multiple subsets, each associates to one city. 


```{r}
# Split climate in multiple dataframe, each associates with one city
cities <- climate %>% 
  group_split(loc)

# Let's examine the structure of variable cities 
str(cities)

# name each elment of "cities" with the correct name
Alert <- cities[[1]]
Mould_Bay <- cities[[2]]
Sachs_Habor <- cities[[3]]
Victoria <- cities[[4]]
Windsor <- cities[[5]]
Yarmouth <- cities[[6]]
```

Second, we will draw multiple temporal-trend plots. One for each city. To do this, I will write a function that input **city result** and output the **scatter plot**. This function would be based heavily on the lab manual provided by professor Julien Martin

```{r}
plotting_temporal_trend <- function (city){ 
  
  #city: Yarmouth, Windsord,...as a data frame

mygraph <- ggplot(
  data = city[!is.na(city$ave), ], # source of data
  aes(x = year, y = ave)
)

# plot data points, regression, loess trace
mygraph <- mygraph +
  stat_smooth(method = lm, se = FALSE, color = "green") + # add linear regression, but no SE shading
  stat_smooth(color = "red", se = FALSE) + # add loess
  geom_point()+
  ggtitle( "Rate of change in temperature of", deparse(substitute(city)))         # add data points

mygraph
}

```


Next, I will draw all 6 temporal trend graphs for all 6 cities

```{r}
plotting_temporal_trend(Alert)
plotting_temporal_trend(Mould_Bay)
plotting_temporal_trend(Sachs_Habor)
plotting_temporal_trend(Victoria)
plotting_temporal_trend(Windsor)
plotting_temporal_trend(Yarmouth)
```


After the temporal trends, we want to estimate the rate of change (or the slope) in temperature for each city. 

I will summarize all linear regression result in one table, so it is easy to compare

```{r}
fitted_models = climate %>% 
                group_by(loc) %>% 
                do(fit_ave = lm(ave ~year, data = .)) %>% 
                ungroup %>% 
                mutate(yearCoef = map(fit_ave, tidy))%>%
                unnest(yearCoef)

fitted_models
```

From the table, we can see the *slope* of each city is the **estimation**. I will display the result in a simpler table down below


```{r}
rate_of_change <-fitted_models %>% filter(term == "year")%>%
                select(loc, estimate) %>% 
                transmute(city= loc, rate_of_change=estimate)

rate_of_change
```



6. Compute the mean warming rate in the North and in the South by averaging the rates calculated separately for each station (1 point). Calculate a 95% CI for each of these means based on the 3 replicates stations. (2 points)




-----------------------

Do not forget if you want to use Rmarkdown, first add multiple empty lines between each questions then add code chunks to type your R code in them. and write your interpretation below.
    
    
    
    
    

# X. example question

```{r}
#write your code here

```

Write your interpretation