---
title: "Assignment_2_BIO_4158"
author: "Michael Pham (300129636), Yacine Marouf (300112014), Tristan Lachance (300059877)"
date: "2022-09-28"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Here are the needed packages
```{r}

### Load the packages needed
library(tidyverse)
library(ggplot2)
library(readr)
library(pwr)
library(tinytex)
```


We will load the dataset to R in here: 

```{r} 
setwd("/Users/apple/Desktop/BIO_4158 /BIO 4158 lab/Data sets BIO 4158")
getwd()
mydata <-read_csv("simulies.csv")

```

#1. Calculate the mean mass and its standard deviation at each site.


```{r} 

mydata_new <-mydata %>% mutate(Mass = 1.36 *(Length^3.05))

mydata_new %>% 
  group_by(Site) %>% 
  summarise(mean= mean(Mass), standard_deviation=sd(Mass))
```



#2. Using the mean mass of P. mixtum/fuscum at Lovering as a reference and the standard deviations observed at each site, calculate the power of a two-tailed t test for independent means (4 points)

We have: u1=13.1, u2=43.0, s1=23.8, s2=48.0

```{r}
# Function to calculate effective size 
d <- function(u1, u2, sd1, sd2) {
  abs(u1 - u2) / sqrt((sd1^2 + sd2^2) / 2)
}

# Power Analysis for 2 population ( not in the assignment )
pwr.t.test(
  n = 100,
  d = d(u1 = 13.1 , sd1 = 23.8, u2 = 43.0, sd2 = 48.0),
  sig.level = 0.05,
  type = "two.sample")

```




2.a) if the mass difference was 5 µ g, α=0.05, and that 100 larvae were sampled at each site

```{r}

pwr.t.test(
  n = 100,
  d = d(u1 = 13.1 , sd1 = 23.8, u2 = 13.1+5 , sd2 = 48.0),
  sig.level = 0.05,
  type = "two.sample")

```


From the output, we can see that the power is **power= 0.153**





2.b) if the mass difference was 20 µ g, α=0.05, and that 100 larvae were sampled at each site

```{r}
pwr.t.test(
  n = 100,
  d = d(u1 = 13.1 , sd1 = 23.8, u2 = 13.1+20 , sd2 = 48.0),
  sig.level = 0.05,
  type = "two.sample")
```

From 2b output, we can see that the power is: **power= 0.96**

2.c) if the mass difference was 50 µ g, α=0.05, and that 100 larvae were sampled at each site




```{r}
pwr.t.test(
  n = 100,
  d = d(u1 = 13.1 , sd1 = 23.8, u2 = 13.1+50 , sd2 = 48.0),
  sig.level = 0.05,
  type = "two.sample")
```
From the output of 2c, we can see that the power is: **power= 1**


2d) As we can see from other part of the question, power is directly proportional to the sample size. The larger the sample size, the larger the power. 



#3. Calculate the required sample size to detect, with a two-tailed t-test for independent means, a difference of 50 µ g between the means (4 points)
        Answer
3.a) with a power of 80% and α=0.05

```{r}
## This piece of code will caculate detectable effect size
pwr.t.test(
  power = 0.8, # Power 
  d = d(u1 = 13.1 , sd1 = 23.8, u2 = 13.1+50 , sd2 = 48.0), ## mean is 50ug
  sig.level = 0.05, 
  type = "two.sample")
```


From the 3a output, we can see that the sample size is **n=11 **




 3.b) with a power of 80% and α=0.001

```{r}
## This piece of code will caculate detectable effect size
pwr.t.test(
  power = 0.8, # Power 
  d = d(u1 = 13.1 , sd1 = 23.8, u2 = 13.1+50 , sd2 = 48.0), ## mean is 50ug
  sig.level = 0.001, 
  type = "two.sample")
```
From the 3b output, we can see that the sample size is **n=23 **

3.c) with a power of 95% and α=0.05

```{r}
## This piece of code will caculate detectable effect size
pwr.t.test(
  power = 0.95, # Power 
  d = d(u1 = 13.1 , sd1 = 23.8, u2 = 13.1+50 , sd2 = 48.0),  ## mean is 50ug
  sig.level = 0.05, 
  type = "two.sample")
```

From the 3c output, we can see that the sample size is **n=16**

3.d) comment on how changing power and α affects the required sample size.



As from the result I collect from the output of R \

- significant level vs sample size: Signficant level is **indirectly proportional** to sample size

- power vs sample size: power is **directly proportional** to sample size


#4. Calculate the detectable effect size (d) with a two-tailed t-test on independent means, given the observed standard deviations (4 points)

4.a) with a power of 80%, α=0.05, and measurements on 10 larvae at each site

```{r}
pwr.t.test(
  power = 0.8, # Power 
  n = 10, ## Sample size 
  sig.level = 0.05, 
  type = "two.sample")

```

From output of 4a, we can see that the detectable effect size is **d=1.324947**



4b) with a power of 80%, α=0.05, and measurements on 200 larvae at each site

```{r}
pwr.t.test(
  power = 0.8, # Power 
  n = 200, ## Sample size 
  sig.level = 0.05, 
  type = "two.sample")

```

From output of 4a, we can see that the detectable effect size is **d = 0.28**

4.c) with a power of 80%, α=0.05, and measurements on 20 larvae at one site and 380 larvae at the second site

```{r}
pwr.t2n.test(
  
  n1=20,
  n2=380,
  power=0.8,
  sig.level=0.05
)

```

From the output of 4c, we can see that the detectable effect size is: **d = 0.6442829**


 4.d) comment on how changing sample sizes affect d.


As we can see from the result, sample size is **indirectly proportional** to the effective size





5. Calculate the difference in average mass, in µ g, that can be detected given your estimate of the minimum detectable effect size (d) at 4a, b, and c. (4 points)


a) Difference in average mass in ug of 4a (power=0.8, alpha=0.05, n=10)


```{r}
pwr.t.test(
  power = 0.8,
  n = 10,
  sig.level = 0.05,
  type = "two.sample")$d * sqrt((23.8^2 + 48.0^2) / 2)
```

From the output, we can see that the difference is: **difference= 50.19467**


b )Difference in avarage mass in ug of 4b (power=0.8, alpha=0.05, n=200)

```{r}

pwr.t.test(
  power = 0.8,
  n = 200,
  sig.level = 0.05,
  type = "two.sample")$d * sqrt((23.8^2 + 48.0^2) / 2)
```

From the output, we can see that the difference is: **difference= 10.63892**




c) Difference in average mass in ug of 4c (power=0.8, alpha=0.05, n1=20,n2=80)

```{r}

pwr.t2n.test(
  
  n1=20,
  n2=380,
  power=0.8,
  sig.level=0.05
)$d * sqrt((23.8^2 + 48.0^2) / 2)

```

From the output, we have the difference is: **difference = 24.40821**

